---
title: "Dating with Data - Exploring OKCupid Profiles to Understand User Characteristics"
author: "Team Extrapolats"
date: "`r Sys.Date()`"
output:  
    rmdformats::readthedown:
      toc_float: true
      toc_depth: 3
      number_sections: false
      code_folding: hide
#  includes:
#        before_body: header.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, results = "show", message = F)
library(ezids)
```

# ABSTRACT

This project presents an analysis of the OKCupid dataset, focusing on the relationships between demographic characteristics, lifestyle choices, and physical attributes of users. The analysis involves multiple steps: data cleaning, exploratory data analysis (EDA), and hypothesis testing. 

Initially, we processed the dataset by converting relevant character columns into factors and removing outliers. We then explored the distribution of key variables such as age, height, gender, and relationship status through descriptive statistics and visualizations. 

Our findings reveal significant insights, including a notable correlation between education levels and drug use, as well as the prevalence of specific body types among social drinkers who smoke. Additionally, we examined the relationship between dietary preferences and body type, finding an association between diet category and healthiness of body types.

Using chi-squared tests, we confirmed the statistical significance of these relationships, allowing us to reject the null hypotheses for both lifestyle factors. Overall, this analysis underscores the complex interplay between personal choices and demographic factors within the context of online dating profiles.


# INTRODUCTION

The rise of online dating platforms, particularly OKCupid, has generated extensive user data that offers insights into relationship dynamics. This project focuses on analyzing the OKCupid dataset to explore the relationships between demographic attributes, lifestyle choices, and physical characteristics of users. By examining factors such as educational attainment, substance use, drinking and smoking habits, and dietary preferences, we aim to uncover patterns that influence romantic connections.

Using statistical methods and visualizations, this analysis seeks to provide a deeper understanding of how these variables interact within the context of online dating. The findings will contribute to broader discussions about social behaviors and health outcomes in contemporary society, emphasizing the relevance of data-driven insights in shaping our understanding of personal relationships in the digital age.

# SUMMARY OF THE DATASET

The original OkCupid dataset was 59,946 observations of 31 variables. The data also included user characteristics: age-dev, which ranged from 22 to 38 in integers;

# DATASET OVERVIEW

The OkCupid dataset contains 59,946 observations on 31 variables about users in the San Francisco Bay Area in 2012. Examples of the wide range of user information captured by data are demographic information, which includes age, sex, status, and orientation; background information, which includes ethnicity, religion, and language proficiency; location information from various Bay Area cities. Physical attributes of height and body_type and lifestyle choices, such as drinks, smokes, drugs, and diet. Personal preferences regarding pets and offspring are also represented. This dataset contains education and career-related information: education level, income, job, timestamps of user activity, and essay responses-essay0-essay2-which include self-descriptions. Although raw data provides a broad range of user characteristics, several data quality issues-which include inconsistencies and missing values-make it hardly analyzable without extensive preprocessing.

## Loading the Dataset 

### Before Cleaning
```{r}
# Reading the dataset
cupid_df = read.csv("OKCUPID.csv")
```

```{r}
# Structure of the dataset
str(cupid_df)
```

# DATA CLEANING

We started by importing the dplyr package, which is useful for data manipulation in R. Then, we concentrated on cleaning up the cupid_df dataset by removing unnecessary columns. To create a list of columns to exclude, we utilized names() and paste0(), focusing on the columns falling between essay0 and essay9. Additional columns such as religion, last_online, income, sign, speaks, pets, and offspring were removed to focus our analysis on the pertinent information. 

```{r}
# Removing unnecessary columns from dataframe 

library(dplyr)

cupid_df <- cupid_df[, !names(cupid_df) %in% paste0("essay", 0:9)]
cupid_df <- cupid_df %>% select(-religion)
cupid_df <- cupid_df %>% select(-last_online)
cupid_df <- cupid_df %>% select(-income)
cupid_df <- cupid_df %>% select(-sign)
cupid_df <- cupid_df %>% select(-speaks)
cupid_df <- cupid_df %>% select(-pets)
cupid_df <- cupid_df %>% select(-offspring)
```

Next, we examined the dataset to identify any instances of empty strings or NA values. We used sapply() to apply a function to all columns in cupid_df to count empty strings and NA values. 
This has allowed us to gain a clear comprehension of the data quality by identifying the amount of empty strings and missing values in each column. 


```{r}

# Check for empty and NA values 
empty_values <- sapply(cupid_df, function(x) sum(x == "", na.rm = TRUE))
na_values <- sapply(cupid_df, function(x) sum(is.na(x)))
print("Empty Values in Each Column:")
print(empty_values)
print("NA Values in Each Column:")
print(na_values)
```

After we found these problems, we substituted all absent strings with NA values. 
This simplified the management of data gaps. We confirmed again to ensure that all blank strings were changed to NA, making our dataset more reliable and consistent. 


```{r}
# Replace empty strings ("") with NA values
cupid_df[cupid_df == ""] <- NA

# Check for empty and NA values again
empty_values <- sapply(cupid_df, function(x) sum(x == "", na.rm = TRUE))
na_values <- sapply(cupid_df, function(x) sum(is.na(x)))
print("Empty Values in Each Column after NA conversion:")
print(empty_values)
print("NA Values in Each Column after NA conversion:")
print(na_values)
```


We used subset() and complete.cases() to create a new dataframe, cleaned_cupid_df, by removing any incomplete cases. 
Removing rows with missing values was an essential step to guarantee the accuracy of our analysis. 
After cleaning, we double-checked for missing values to confirm that all rows were fully filled. 



```{r}
cleaned_cupid_df <- subset(cupid_df, complete.cases(cupid_df))
na_values <- sapply(cleaned_cupid_df, function(x) sum(is.na(x)))
print("NA Values in Each Column after cleaning:")
print(na_values)
```


Following that, we provided further clarification on the gender variable. 
Through the use of mutate() and recode(), we converted the abbreviated terms ('f' for female and 'm' for male) to complete words, which enhances the readability and comprehension of the data for future users.



```{r}
# Converting gender shortcut
cleaned_cupid_df <- cleaned_cupid_df %>% mutate(sex = recode(sex, 'f' = 'female', 'm' = 'male'))

# Converting required 'char' columns into 'factor'

cleaned_cupid_df$status <- as.factor(cleaned_cupid_df$status)
cleaned_cupid_df$sex <- as.factor(cleaned_cupid_df$sex)
cleaned_cupid_df$orientation <- as.factor(cleaned_cupid_df$orientation)
cleaned_cupid_df$body_type <- as.factor(cleaned_cupid_df$body_type)
cleaned_cupid_df$drugs <- as.factor(cleaned_cupid_df$drugs)
cleaned_cupid_df$diet <- as.factor(cleaned_cupid_df$diet)
cleaned_cupid_df$drinks <- as.factor(cleaned_cupid_df$drinks)
cleaned_cupid_df$education <- as.factor(cleaned_cupid_df$education)
cleaned_cupid_df$ethnicity <- as.factor(cleaned_cupid_df$ethnicity)
cleaned_cupid_df$job <- as.factor(cleaned_cupid_df$job)
cleaned_cupid_df$smokes <- as.factor(cleaned_cupid_df$smokes)
cleaned_cupid_df$smokes <- as.factor(cleaned_cupid_df$smokes)
```

Next, we converted multiple character columns into factors using the as.factor() function. 
This step is essential when working with categorical data in R because it helps with organization and allows for more accurate statistical analysis. 
Altering these columns was crucial for any forthcoming modeling we intended to conduct. 


## Summary of the cleaned Dataset
```{r}
summary(cleaned_cupid_df)
```

After we examined the data, we condensed it with the summary() function. 
This enabled us to develop a thorough comprehension of the dataset, aiding us in understanding its distribution and characteristics before diving into a more in-depth analysis.



# EXPLORATORY DATA ANALYSIS

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(psych)
library(tidyverse) 
library(ggplot2) 
library(viridis) 
library(RColorBrewer) 
library(ggmosaic)
```

## Distribution of Numeric Variables

```{r}
# Descriptive Statistics
# Numeric variables
describe(cleaned_cupid_df %>% select(age, height))
```

We started our data exploration by importing key libraries such as ggplot2, dplyr, and other R tools to assist us in the efficient analysis and visualization of the data. 

Our primary goal was to examine how quantitative variables, like age and height, were distributed through the use of descriptive statistics. This helped us gain a better understanding of the data. 
We used the describe() function to produce a summary of the variables in the cleaned_cupid_df dataset.


### Age

The  code generates a histogram to visualize the age distribution of users in the OKCupid dataset. Here's a breakdown of the key components and what they reveal:
Data Input:
The ggplot function initializes the plotting system using the cleaned dataset cleaned_cupid_df, with the age variable mapped to the x-axis.

Histogram Creation:
The geom_histogram function creates the histogram. The binwidth parameter is set to 5, meaning ages will be grouped into 5-year intervals. This choice balances detail with clarity, allowing for an effective representation of age distribution without overwhelming the viewer with too many bins.
The fill argument assigns a blue color to the bars, while color = "black" outlines the bars, enhancing visual contrast and making the distribution clearer.

```{r}
# Graphical Representations

# Histogram of age
ggplot(cleaned_cupid_df, aes(x = age, fill = age)) + geom_histogram(binwidth = 5, fill = "blue", color = "black") + labs(title = "Age Distribution", x = "Age", y = "Frequency") + theme_minimal()
```
Labels and Themes:
labs is used to add titles and axis labels, which provide context for the graph. The title "Age Distribution" briefly summarizes what the viewer is examining, while the x-axis and y-axis labels indicate the variables represented.
The theme_minimal() function applies a clean and simple aesthetic to the plot, minimizing distractions and focusing attention on the data.
Insights and Interpretation:
By analyzing the histogram, viewers can identify trends such as the most common age ranges among users, any skewness in the distribution, and the overall age demographic of the user base.

For instance, if there are noticeable peaks at certain ages, this could indicate popular age groups on the platform. Conversely, if the distribution is flat, it may suggest a more diverse age representation.
Overall, this histogram serves as a foundational analysis tool, providing essential insights into the age demographics of the OKCupid user base that can inform subsequent analyses related to user behaviors and preferences.
4o mini

### Height
```{r}
# Density plot for height
ggplot(cleaned_cupid_df, aes(x=height, fill = height)) +  geom_bar(fill="green", color="black")+ labs(title = "Height Distribution", x = "Height", y = "Frequency") + theme_minimal()
```

Next, we made a density plot to show how heights are spread out among different body types. We used geom_bar to display different height values, filling the bars with green and then outlining them in black. 
The plot included titles and markings on the axes for clearer comprehension. 


### Outlier removal
```{r}
cleaned_cupid_df <- outlierKD2(cleaned_cupid_df, height, rm = TRUE, boxplt = TRUE, histogram = FALSE, qqplt = TRUE)
```

### Gender Distribution
```{r}
ggplot(cleaned_cupid_df, aes(x=sex, fill=sex))+geom_bar(color = "black")+ggtitle("Male and Female count in Dataset") + theme_minimal()+ xlab("Gender") +  ylab("Frequency")
```

Subsequently, we proceeded to investigate the gender distribution in the dataset. A bar graph was made using ggplot to display the numbers of males and females, filled with corresponding colors and bordered in black. 
The plot was titled "Count of Men and Women in Dataset," providing a clear overview of gender distribution. 


### Marital Status
```{r}
# Bar plot for status
ggplot(cleaned_cupid_df, aes(x = status, fill = status)) + geom_bar(fill = "orange") + labs(title = "Status Distribution (Beside Single)", x = "Status", y = "Count") + theme_minimal() + geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3)
```

Next, we analyzed marital status by utilizing an alternate bar graph. The bars were colored orange to indicate the quantities of various status groups in the visualization. 
Labels were included on the bars to display the quantity for every category, enhancing the data's clarity. 


### Distribution of Height by Body Type
```{r}
# Boxplot of height by body type
ggplot(cleaned_cupid_df, aes(x = body_type, y = height)) + geom_boxplot(fill = "lightblue") + labs(title = "Height by Body Type", x = "Body Type", y = "Height") + theme_minimal()

```

Lastly, we looked at the relationship between height and body shape by using a boxplot. The chart displayed the relationship between height and body shape, using light blue shaded boxes. 
These visuals provided valuable insights into the data by showing the varying heights among different body types. 


Overall, the use of both graphs and numerical summaries helped us comprehend the data's structure and the relationships between various variables.


# SMART QUESTIONS

### Question 1: What is the distribution of ages among users, and how does it differ between genders?

```{r}
# scatter plot
ggplot(cleaned_cupid_df, aes(x = age, color = sex)) + geom_freqpoly(stat = "count", aes(y = ..count..)) + labs(title = "Age Distribution by Gender", x = "Age", y = "Number of Users") + scale_color_manual(values = c("blue", "red"), labels = c("Female", "Male")) + theme_minimal()

```
The Scatter plot above shows Age distribution by gender indicates that Males aged 25-26 and Females around 25-28 are highest in number. The majority of users are males, but after age 55, females are more.

### Question 2: Which category of educated people are more inclined towards often taking drugs?

For this, first we group people who takes drugs 'often' and 'sometimes' as drug users.
```{r}
# Recode and filter dataset
frequent_drug_users_df <- cleaned_cupid_df %>%mutate(drugs = ifelse(drugs %in% c("sometimes", "often"), "Drugs", drugs)) %>% filter(drugs == "Drugs")

# Bar plot 
ggplot(frequent_drug_users_df, aes(x = education, fill = drugs)) +geom_bar(position = "dodge") +labs(title = "Education Level Among Frequent Drug Users", x = "Education Level",y = "Count of Individuals",fill = "Drug Usage"
)+scale_fill_manual(values = "red") + theme_minimal() +theme(axis.text.x = element_text(angle = 90, hjust = 1), plot.title = element_text(hjust = 0.4, size = 16, face = "bold")) +geom_text(stat = "count", aes(label = ..count..), position = position_dodge(width = 0.9), vjust = -0.3, size = 2)
```

Our analysis of the pattern of drug usage by education level is quite illustrative. As clearly denoted in the visualization, the group that reported completing college/university had the highest disposition towards the use of drugs, with 1,641 users. This then follows those currently working towards their college/university education, with 685 users. Of interest, the most conservative patterns of drug use were found among those subjects who had dropped out from colleges, universities, and law schools. Further analysis may thus reveal that completion of higher education, especially college/university level, is positively correlated with higher reported drug use, while discontinuation of education results in more conservative usage. Conclusively, these findings give an excellent insight into the association of drug use behavior with educational attainment in the dataset of dating profiles.

### Question 3: Does the drinking and smoking habit of an individual who drinks 'socially'and smokes definitively explain the body type of them?

In order to understand how the body type of individuals who both drink socially and smoke can be influenced by their drinking and smoking habits, we initially narrowed down our data to this particular group, forming a subset known as social_smokers_df. This specific group consists of individuals who engage in social drinking and smoking, but does not include non-smokers. 

We generated a bar chart to display how body types are distributed among these occasional smokers in our study. The storyline uncovers a few fascinating patterns: most people in this category have "average," "fit," and "athletic" body shapes. On the other hand, there is a distinct decrease in the number of people labeled as "ripped," "obese," or "spent." This pattern indicates that casual drinkers who also smoke tend to have a more average or healthier body shape compared to individuals with more drastic body shapes. 

```{r}
# filter dataset
social_smokers_df <- cleaned_cupid_df %>% filter(drinks == "socially" & smokes != "no")

# Bar plot
ggplot(social_smokers_df, aes(x = body_type, fill = body_type)) + geom_bar() + scale_fill_brewer(palette = "Set3") + 
labs(title = "Body Type Distribution Among Social Drinkers Who Smoke", x = "Body Type", y = "Number of Individuals") + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3)

```

Most of the people who drink 'socially' and smoke are average, fit and even athletic. People with body types jacked, overweight and used up are lowest in number. 

```{r}
# Chi-square test
chi_square_result <- chisq.test(table(social_smokers_df$body_type))
print(chi_square_result)
```

Next, We conducted a Chi-squared test to examine the relationship between body types and drinking behaviors. This test is appropriate for this situation since we are working with classified data. The p-value obtained from the Chi-squared test was less than 0.05. This indicates that we can reject the null hypothesis that states there is no correlation between drinking habits and body types.

The results reveal an interesting link between how people drink socially and their body types. It seems that body shapes might be connected to casual drinking and smoking, possibly influenced by lifestyle and social settings. Overall, our findings point to the need for more research into how these habits could affect someone’s appearance.



### Question 4: Is there a significant association between an individual’s diet preference (ranging from 'anything' to 'strictly vegetarian') and having a 'healthy' body type?

```{r}
# Filtering the data
diet_body_df <- cleaned_cupid_df %>%filter(diet %in% c("anything", "strictly anything", "mostly vegetarian", "strictly vegetarian")) %>%mutate(diet_category = ifelse(diet %in% c("anything", "strictly anything"), "Anything", "Vegetarian"),healthy_body_type = ifelse(body_type %in% c("athletic", "average", "fit", "jacked"), "Healthy", "Unhealthy"))

# bar plot 
ggplot(diet_body_df, aes(x = diet_category, fill = healthy_body_type)) +geom_bar(position = "dodge") +labs(title = "Diet Preference and 'Healthy' Body Type Distribution",x = "Diet Category",y = "Number of Individuals",fill = "Body Type (Healthy)") +scale_fill_brewer(palette = "Set3") +theme_minimal() +geom_text(stat = "count", aes(label = ..count..),position = position_dodge(width = 0.9), vjust = -0.5, size = 3) 

```

people who eat both vegetarian and non-vegetarian were more compared to people who stick with vegetarian food. But the difference in healthy and unhealthy body is less for vegetarians compared to people who eat anything.

```{r}
# Chi-square
diet_body_table <- table(diet_body_df$diet_category, diet_body_df$healthy_body_type)
chi_square_result <- chisq.test(diet_body_table)
print(chi_square_result)
```

We perform Chi-squared test because the variables we work on are categorized. We can see the p-value is equal to 0.05. So we can reject the Null hypothesis and can conclude there is a association between diet inntake and body type a person has.

# CONCLUSION



# REFERENCES

1. Sindy R Sumter, Laura Vandenbosch, Dating gone mobile: Demographic and personality-based correlates of using smartphone-based dating applications among emerging adults.
2. [W3 Schools](https://www.w3schools.com/r/r_graph_plot.asp)
