---
title: "PREDICTING THE PROPOSAL: A Data-Driven Approach to Love"
author: "Maddala Harshith"
date: "`r Sys.Date()`"
output:  
    rmdformats::readthedown:
      toc_float: true
      toc_depth: 3
      number_sections: false
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, results = "show", message = F)
library(ezids)
```

# ABSTRACT
```{r}
```

# INTRODUCTION
```{r}
```

# SUMMARY OF THE DATASET
```{r}
```

# DATASET OVERVIEW
```{r}
```

## Loading the Dataset 
```{r}
```

### Before Cleaning
```{r}
```

# DATA CLEANING
```{r}
```

## Summary of the cleaned Dataset
```{r}
```

# EXPLORATORY DATA ANALYSIS
### Loading necessary libraries

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(psych)
library(tidyverse) 
library(ggplot2) 
library(viridis) 
library(RColorBrewer) 
library(ggmosaic)

library(gridExtra)
library(caret)
library(pROC)
library(rpart)
library(rpart.plot)
library(randomForest)
```


### Distribution of the Numeric variables

Let’s take a look at descriptive stats of Numeric variables.
```{r}
# Descriptive Statistics
# Numeric variables
describe(proposal_df %>% select(Age, Height, Income, DistanceKM))
```

Graphical Representations of Integer variables.

```{r}
# Histogram of age

ggplot(proposal_df, aes(x = Age, fill = Age)) + geom_histogram(binwidth = 5, fill = "lightpink", color = "black") + labs(title = "Age Distribution", x = "Age", y = "Frequency") + theme_minimal()
```

```{r, warning=FALSE}
# Frequency polygon for Height
ggplot(proposal_df, aes(x = Height)) + 
  geom_freqpoly(binwidth = 5, color = "grey", size = 1.2) + 
  labs(title = "Height Distribution", x = "Height (cm)", y = "Frequency") + 
  theme_minimal()

```


```{r}
# We can use this function to check if any outliers were detected in variables.
#proposal_df <- outlierKD2(proposal_df, DistanceKM, rm = TRUE, boxplt = TRUE, histogram = FALSE, qqplt = TRUE)
```

```{r}
ggplot(proposal_df, aes(x = cut(Income, breaks = c(5000, 20000, 30000, 40000, 50000), labels = c("Low Income", "Medium Income", "High Income", "Very High Income"), include.lowest = TRUE))) + geom_bar(fill = "steelblue", color = "black") + labs(title = "Population Distribution by Income Category", x = "Income Category", y = "Frequency") + theme_minimal()

```

```{r}
# Boxplot for Distance
ggplot(proposal_df, aes(y = DistanceKM)) + geom_boxplot(fill = "lightyellow", color = "black") + labs(title = "Distance (KM) Boxplot", y = "Distance (KM)") + theme_minimal()
```

Graphical Representations of Factorial variables.

```{r, warning=FALSE}
# Pie chart for Response variable
ggplot(proposal_df, aes(x = "", fill = Response)) + geom_bar(width = 1, stat = "count", color = "black") + coord_polar(theta = "y") + geom_text(aes(label = scales::percent(..count../sum(..count..))), position = position_stack(vjust = 0.5), stat = "count") + scale_fill_manual(values = c("0" = "red", "1" = "green"), labels = c("Rejected", "Accepted")) + labs(title = "Proposal Response Distribution", fill = "Response") + theme_void() + theme(plot.title = element_text(hjust = 0.5, size = 16))
```

```{r}
# bar plot for Communication Score
p1 <- ggplot(proposal_df, aes(x = as.factor(CommunicationScore))) + geom_bar(fill = "blue", color = "black") + labs(title = "Communication Score Distribution", x = "Communication Score", y = "Frequency") + theme_minimal()

# bar plot for Compatibility Score
p2 <- ggplot(proposal_df, aes(x = as.factor(CompatibilityScore))) + geom_bar(fill = "green", color = "black") + labs(title = "Compatibility Score Distribution", x = "Compatibility Score", y = "Frequency") + theme_minimal()

grid.arrange(p1, p2, ncol = 2)
```

## Distribution of Numeric Variables
```{r}
```

### Outlier removal
```{r}
```

# SMART QUESTIONS
```{r}
```

### Question 1:
```{r}
```

#### OBSERVATION:
```{r}
```

#### RESULT:
```{r}
```

### Question 2:
```{r}
```

#### OBSERVATION:
```{r}
```

#### RESULT:
```{r}
```

### Question 3:
```{r}
```

#### OBSERVATION:
```{r}
```

#### RESULT:
```{r}
```

### Question 4:
```{r}
```

#### OBSERVATION:
```{r}
```

#### RESULT:
```{r}
```

# CONCLUSION
```{r}
```

# REFERENCES
```{r}
```



