---
title: "PREDICTING THE PROPOSAL: A Data-Driven Approach to Love"
author: "Maddala Harshith"
date: "`r Sys.Date()`"
output:  
    rmdformats::readthedown:
      toc_float: true
      toc_depth: 3
      number_sections: false
      code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, results = "show", message = F)
library(ezids)
```

# ABSTRACT
```{r}
```

# INTRODUCTION
```{r}
```

# SUMMARY OF THE DATASET
```{r}
```

# DATASET OVERVIEW
```{r}
```

## Loading the Dataset 
```{r}
```

### Before Cleaning
```{r}
```

# DATA CLEANING
```{r}
```

## Summary of the cleaned Dataset
```{r}
```

# EXPLORATORY DATA ANALYSIS
```{r}
```

## Distribution of Numeric Variables
```{r}
```

### Outlier removal
```{r}
```

# SMART QUESTIONS
```{r}
```

### Question 1:
```{r}
```

#### OBSERVATION:
```{r}
```

#### RESULT:
```{r}
```

### Question 2:
```{r}
```

#### OBSERVATION:
```{r}
```

#### RESULT:
```{r}
```

### Question 3:
```{r}
```

#### OBSERVATION:
```{r}
```

#### RESULT:
```{r}
```

### Question 4:
```{r}
```

#### OBSERVATION:
```{r}
```

#### RESULT:
```{r}
```

# CONCLUSION
```{r}
```

# REFERENCES
```{r}
```



